Open-source landscape for 3D CSEM modelling

### 2020-10-23 manuscript/Manuscript.tex
Submission version

\title[Open-source 3D CSEM modelling]{Open-source landscape for 3D CSEM modelling}

@@

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most commonly applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{GEO.93.Wang, RS.94.Druskin, RSC.94.Mackie, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or octree meshes \citep{ECP.07.Haber}.

@@

Today it is becoming more common for researchers in many domains of science to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an \emph{open model of development} where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of (open-source) code towards building global communities around open projects. Other notable Python projects with an open model of development within the same realm as the codes under considerations are \texttt{pyGIMLi} \citep{CAG.17.Rucker}, \texttt{Fatiando} \citep{JOSS.18.Uieda}, \texttt{GemPy} \citep{GMD.19.DeLaVarga}, and \texttt{PyVista} \citep{JOSS.19.Sullivan}.

@@

We briefly present our codes in this section. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the codes is given in Table~\ref{tbl:codecomparison}. All four have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range usually applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator. A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations \emph{PEC} and \emph{PMC}, which stand for perfect electric conductor
We could add maths here; but we would need to introduce all the symbols. For now, we leave it out.
% ($\vec{n}\times\vec{E} = 0; \vec{n}\cdot\vec{H} = 0$)
and perfect magnetic conductor%
% ($\vec{n}\cdot\vec{E} = 0; \vec{n}\times\vec{H} = 0$)
, respectively. Another abbreviation used in the tables is \emph{dof} for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real-world time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation time or CPU time is therefore much higher than the reported runtime for the codes that run in parallel, with an upper limit of \#Procs$\times$Runtime. Memory refers to the maximum memory increase at any point of solving the SLE.

@@

For the parallel forward modelling computations, a highly scalable MPI (message passing interface) domain decomposition allows reducing runtimes and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters (e.g., High-Performance Computing (HPC) architectures).

@@

Note that \emg3d and \simpeg used the same rectilinear mesh in this example. The use of an octree mesh in \simpeg could greatly reduce the size of the mesh and resultant computation time, but at the cost of accuracy, which is reduced when averaging over layer interfaces to coarsen cells and because of the factor-of-two expansion rate of coarsened cells. As the aim of this study is to focus on accuracy of the simulations and compare discretization approaches, we chose to use the same mesh as in the \emg3d simulation.
%
The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}
\label{tbl:comp-layered}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 24 &   118 &  97.8 & 1531824 \\ % dof: 3063648 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   116 &   0.4 & 4813216 \\
  \petgem & 24 &   252 & 151.7 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

There are no semi-analytical solutions for such a model to verify our results, and we can only validate them by comparing different results. For this we use the normalised root-mean square difference (NRMSD) between two responses $R_1$ and $R_2$, given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200 \frac{|R_1 - R_2|}{|R_1| + |R_2|}\ .
  \label{eq:nrmsd}
\end{equation}
%

The results for the three receiver lines $y=-3;0;+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the NRMSDs between the absolute responses of the codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-block}
  \caption{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the NRMSDs (\%) between the absolute responses $|E_x|$ of the different codes are shown in the bottom row.}
  \label{fig:results-block}
\end{figure*}
%

A few interesting points stand out:
\begin{enumerate}
  \item The error levels between the different codes are overall comparable; the NRMSD is generally below a few percent, increasing towards larger offsets.
  \item The NRMSD between the FV codes \emg3d and \simpeg is comparably low except at the boundary. The generally low NRMSD is because these codes use the same rectilinear mesh for modelling. The increasing NRMSD towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition in this example.
  \item The FE codes \custem and \petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
  \item The agreement within the same differential equation method (FE or FV) is better than across methods.
\end{enumerate}

@@

The responses for all six frequencies and all three electric components are shown in Figure~\ref{fig:results-marlim-responses}, both inline and broadside. The published responses are shown in colour with markers for the frequency, and the responses from our codes are shown beneath in grey colours in this overview plot.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim-responses}
  \caption{MR3D comparison between our codes (grey lines) and the published data (coloured lines). The grey lines under the coloured lines are not visible in most areas, meaning that they are very similar. However, there are three notable zones, (1) to (3), which are explained further in the text.}
  \label{fig:results-marlim-responses}
\end{figure*}
%
Two CSEM data-sets were published with MR3D, one containing clean responses, and one responses where realistic noise was added. For the comparison we use the clean data, but we indicate the chosen noise level by the horizontal line at \num{2e-15}\,V/m. It can be seen that the grey lines are not visible for most parts, which means that the data agree well. However, there are a few notable points where this is not the case, and they are marked with numbers in the figure.
\begin{enumerate}
  \item There are some noticeable differences for positive offsets of the broadside $E_y$ and $E_z$ components, which are likely related to the bathymetry. The reason why we cannot reproduce the published results exactly is that the code is not open-source, and we do not exactly know what it does internally. See the discussion around Figure~\ref{fig:results-marlim_2published}.
  \item The three highest frequencies become noisy at large offsets for all of our codes (well below any real-world noise level). The reason why the published responses are not noisy is probably because they were computed in the time domain, and transformed to the frequency domain. This explains their smooth behaviour, without saying anything about their accuracy.
  \item The responses for the $E_y$ component of the inline acquisition line do not agree at all. In the 1D case, the inline $E_y$ component would be zero, and the only response we can measure here comes from 3D effects. These responses are very low, roughly two orders of magnitude lower than the $E_x$ component. Not the published nor our responses are stable, and any of the responses is as bad as the other. Tiniest differences in meshing or different interpolation algorithms will have a huge effect here.
\end{enumerate}

It is important to note that the published results nor any of our results is the \emph{thruth}, but the closer they agree with each other the more certain we can be of the outcomes. Figure~\ref{fig:results-marlim_2published} shows the NRMSDs between the published results and our codes for the three strongest components, the inline $E_x$ and the broadside $E_x$ and $E_y$ components, for three frequencies.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2published}
  \caption{NRMSDs for all four codes in comparison with the published data, for the inline $E_x$ field and the broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively. Shown are the three frequencies 0.125\,Hz, 0.5\,Hz, and 1.0\,Hz in the top, middle, and bottom row, respectively.}
  \label{fig:results-marlim_2published}
\end{figure*}
%
A few conclusions can be drawn from this figure: In general the NRMSD is roughly 10\,\% or less. The structured meshes have in some cases a smaller NRMSD, e.g., for negative offsets in the inline and broadside $E_x$ responses, and in other cases the unstructured tetrahedral meshes, e.g., for positive offsets for all broadside $E_y$ responses, and often they have a comparable NRMSD. Some NRMSDs have an interesting step-pattern, which is more pronounced for the structured meshes than for the unstructured meshes; this is related to the annotated point (3) in Figure~\ref{fig:model-marlim}. The origin of this zigg-zagg pattern is most likely the bathymetry. We suspect that the code of the published data did probably use the bathymetry information in addition to the resistivity cube. This would explain why the codes using the tetrahedra mesh, which uses the bathymetry as well, does not show this feature.

The importance of the meshing to the result can be seen in Figure~\ref{fig:results-marlim_2ours}, where we compute the NRMSDs between our codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2ours}
  \caption{NRMSDs, just as in Figure~\ref{fig:results-marlim_2published}, but comparing some of the four codes with each other. \custem and \petgem produce very similar results, which is due to the fact that they use the same mesh. This also shows that the biggest impact in 3D forward modelling comes from the discretization.}
  \label{fig:results-marlim_2ours}
\end{figure*}
%
The most obvious result in that figure is that \custem and \petgem produce results that are almost identical, their NRMSD is generally below 0.03\,\%. The reason is simple: \custem and \petgem use the exactly same mesh for the computation of this example, the only difference is therefore the solver. This exemplifies that the biggest difference between different codes is the discretization; the solver is important when it comes to runtime and memory, but not for the accuracy of the result. As these two are so similar we compare \emg3d and \simpeg only to \custem in Figure~\ref{fig:results-marlim_2ours}, as the comparisons to \petgem would look the same. But in general the NRMSDs between our codes look similar as in the comparison with the published results, and the principal source of the difference must be in the different discretizations.

The actual differences in meshing is shown in the next two figures. Figure~\ref{fig:results-marlim_big} shows on the top row the rectilinear mesh of \emg3d, on the middle row the octree mesh of \simpeg, and on the bottom row the tetrahedral mesh of \custem and \petgem. The actual mesh and the broadside $|E_x|$ and $|E_y|$ fields are shown in the left, middle, and right columns, respectively. The extent of the plot includes the entire computational domain for the tetrahedral mesh (in x and z direction); the rectilinear and octree mesh extents are larger than what is shown (rectilinear: $x=327$ to 454\,km, $z=-57$ to 66\,km; octree: $x=288$ to 492\,km, $z=-123$ to 41\,km).
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_big}
  \caption{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively. The shown dimensions shows the entire computational domain for the tetrahedra mesh, whereas the entire rectilinear and octree meshes are bigger.}
  \label{fig:results-marlim_big}
\end{figure*}
%
This figure is interesting for various points, and the differences between the three meshes are very obvious. Particularly visible is the advantage of octree and tetrahedra to create large cells outside of the region of interest to push away the boundary, whereas a rectilinear grid, even though stretched, has to compute many cells far away of the zone of interest. The different boundary condition also play a part, where the top row used PEC, and the other two rows PMC.

Figure~\ref{fig:results-marlim_survey} shows a zoom-in to the survey domain, the domain of interest, which shows how the different codes discretized the zones of interest.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_survey}
  \caption{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively (3$\times$ vertical exaggeration).}
  \label{fig:results-marlim_survey}
\end{figure*}
%
The most striking point of these mesh comparison plots is that, while the overall fields are similar and the responses at receiver locations agree well, the fields can vary quite a lot in different parts of the model. Two kilometres below the seafloor, as an example, there are already significant changes visible, purely due to meshing. This is an important insight that has to be taken into account when interpreting inversion results.

@@

%% 1. Results: good agreement %%
Overall, we observed an excellent match between the solutions. The amplitudes of the layered model have mostly a relative error of less than 1--2\,\%. Looking individually at the real or imaginary part results in much higher error values, but those are related to the zero crossings of the responses. The cross-validation of the results for the block model yields a similar picture, with NRMSD values within a few percents. For the large, complex MR3D model it is a bit different. First it can be noted that the misfits are higher for higher frequencies. For the lowest frequency, 0.125\,Hz, the NRMSD is a few percent for inline and broadside $E_x$ components, and up to 10\,\% for the broadside $E_y$ component. For higher frequencies these numbers rise to the order of $\pm10\,$\%. The reason for the higher misfit for MR3D than for the layered or block model lies mostly in the discretization, boundary effects, and interpolation strategy. While one might argue that 5--10\,\% NRMSD is a lot, we argue that this is probably as good as it gets, given that the code with which the responses we compare to were computed is not open-source. This has been shown within the process of this study. Our first attempt were based on the originally published, fine MR3D model. This model is very detailed, but its extent is too small in x- and y-directions for CSEM computations, and each of our codes did its own extrapolation. However, none of the codes were able to reproduce the published results. The reason is that there are many ways how you can extrapolate a model. We reached out to the authors of MR3D, and they kindly agreed to also published the upscaled and extended computational model. It is only with this model that we were able to get comparable results.

%% 2. Discretization %%
One of the key insights we gained from this study is that creating appropriate meshes is a difficult and time-consuming task. While there exist now several open-source 3D CSEM codes the capability for automatic meshing lacks behind and currently remains a largely manual task requiring experience in the field. It is important to state that the chosen models favour structured meshes. The first two examples can easily and accurately be represented by regular meshes. Using, for instance, a dipping layer instead of three resistive cubes would turn it around and make the model more favourable for unstructured meshes. MR3D would be an ideal scenario for unstructured meshes, as it is based on also published irregular lithological horizons. Since the corresponding resistivity model, however, was defined on a regular grid, we were forced to approximate the comparatively fine regular discretization by an unstructured one and interpolate the resistivity data for being able to apply the FE codes to this problem. We are completely aware about the ineptness of this re-approximation procedure, unless it served for the cross-validation purposes. Both cases show that automatic and adaptive meshing capabilities are needed, which will increase the flexibility of 3D CSEM modelling. Such meshing codes should also take into account the physics of CSEM computation with its diffusive behaviour. It is also important to state that the unstructured mesh used for the MR3D model incorporates the entire domain and could be used for all source and receiver positions, whereas the rectilinear and octree meshes were designed for the shown receiver and corresponding source lines.

@@

All files to rerun the different models with the four codes and reproduce the shown results are available at Zenodo (DOI) \emph{with the final submission}. For the review process we refer to the github page \href{https://github.com/swung-research/3d-csem-open-source-landscape}{github.com/swung-research/3d-csem-open-source-landscape}.

>>>>

\title[3D CSEM Open-Source Landscape]{Open-source landscape for 3D CSEM modelling}

@@

The continuous improvement of computing power and the CSEM boom in the early 2000's in the hydrocarbon industry led to a wealth of developed numerical solutions and according publications. The most commonly applied methods to solve Maxwell's equation are the IE method \citep{GJI.74.Raiche, RS.02.Hursan, GEO.06.Zhdanov, GP.10.Tehrani, CAG.16.Kruglyakov, MGS.17.Kruglyakov} and different variations of the differential equation (DE) method, such as finite differences (FD) \citep{GEO.93.Wang, RS.94.Druskin, RSC.94.Mackie, GEO.09.Streich, CAG.13.Sommer}, finite elements (FE) \citep{GEO.04.Commer, GJI.11.Schwarzbach,GEO.12.daSilva, GJI.13.Grayver, GJI.13.Puzyrev, SEG.16.Zhang}, and finite volumes (FV) \citep{EM.90.Madsen, SIAM.01.Haber, PIER.01.Clemens, GEO.14.Jahandari}. There are also many different types of discretization, where the most common ones are regular grids (Cartesian, rectilinear, curvilinear), mostly using a Yee grid \citep{IEEE.66.Yee} or a Lebedev grid \citep{CMMP.64.Lebedev}, but also unstructured tetrahedral grids \citep{SEG.16.Zhang, CAG.17.Cai}, hexagonal meshes \citep{CAG.14.Cai}, or octree meshes \citep{ECP.07.Haber}.

@@

Today it is becoming more common for researchers in many domains of science to release source-code with an open license that allows use and modification (e.g., see the Open Source Initiative approved licenses: \href{https://opensource.org/licenses}{opensource.org/licenses}). This is an important step for improving reproducibility of research, «\emph{to provide the means by which the reader can verify the validity of the results and make use of them in further research}» \citep{GEO.17.Broggini}. Going a step beyond releasing open-source software, many groups adopt an \emph{open model of development} where code is hosted and versioned in an online repository, all changes are public, and users can engage by submitting and track issues. Community oriented projects further engage by encouraging pull-requests, which are suggested changes to the code. Successful, well-maintained projects often have unit-testing and continuous integration that runs those tests with any changes to the code. Additionally, documentation that includes examples and tutorials is an important component for on-boarding new users and contributors. As a result, in many areas of the geosciences, we are seeing a shift away from a one-way distribution of (open-source) code towards building global communities around open projects. Other notable Python projects with an open model of development within the same realm as the codes under consideration are \texttt{pyGIMLi} \citep{CAG.17.Rucker}, \texttt{Fatiando} \citep{JOSS.18.Uieda}, \texttt{GemPy} \citep{GMD.19.DeLaVarga}, and \texttt{PyVista} \citep{JOSS.19.Sullivan}.

@@

We briefly present our codes in this section. It is, however, beyond the scope of this article to go into every detail of the different modellers, and we refer to their documentations for more details. An overview comparison of the codes is given in Table~\ref{tbl:codecomparison}. All four have in common that they solve Maxwell's equation in its differential form (DE) under the quasi-static or diffusive approximation, hence neglecting displacement currents, which is a common approximation for the low frequency range usually applied in CSEM. The machines on which the different codes were run are listed in Table~\ref{tbl:machines} together with the responsible operator. A few clarifying words on abbreviations and definitions: For the boundary conditions (BC) in the comparison table we use the abbreviations \emph{PEC} and \emph{PMC}, which stand for perfect electric conductor
% We could add maths here; but we would need to introduce all the symbols. For now, we leave it out.
% ($\vec{n}\times\vec{E} = 0; \vec{n}\cdot\vec{H} = 0$)
and perfect magnetic conductor%
% ($\vec{n}\cdot\vec{E} = 0; \vec{n}\times\vec{H} = 0$)
, respectively. Another abbreviation used in the tables is \emph{dof} for degree of freedom, which is equivalent to the size of the SLE we are solving. Finally, we use \emph{runtime} as the wall time or elapsed real-world time from start to end of solving the SLE; pre- and post-processing (e.g., mesh generation) is not measured. Note that the actual computation time or CPU time is therefore much higher than the reported runtime for the codes that run in parallel, with an upper limit of \#Procs$\times$Runtime. Memory refers to the maximum memory increase at any point of solving the SLE.

@@

For the parallel forward modelling computations, a highly scalable MPI (message passing interface) domain decomposition allows reducing runtimes and the solution of large-scale modelling cases. This strategy is capable of exploiting the parallelism offered by both modest multi-core computers and cutting-edge clusters (e.g., High-Performance Computing [HPC] architectures).

@@

Note that \emg3d and \simpeg used the same rectilinear mesh in this example. The use of an octree mesh in \simpeg could greatly reduce the size of the mesh and resultant computation time, but at the cost of accuracy, which is reduced when averaging over layer interfaces to coarsen cells and because of the factor-of-two expansion rate of coarsened cells. As the aim of this study is to focus on accuracy of the simulations and compare discretization approaches, we chose to use the same mesh as in the \emg3d simulation.

The runtime and memory requirements for the layered model are listed in Table~\ref{tbl:comp-layered}.
%
\begin{table*}
\begin{minipage}{10cm}
  \centering
  \caption{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the layered model.}
\label{tbl:comp-layered}
  \begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
  \toprule
  Code & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  \custem & 24 &   118 &  97.8 & 1531824 \\ % dof: 3063648 / 2 (Real+Imag -> Complex)
  \emg3d  &  1 &   116 &   0.4 & 4813216 \\
  \petgem & 24 &   252 & 151.7 & 2455868 \\
  \simpeg &  4 & 10089 & 280.3 & 4813216 \\ % nprocs: 1*4 threads
  \bottomrule
\end{tabular}
\end{minipage}
\end{table*}
%

@@

There are no semi-analytical solutions for such a model to verify our results, and we can only validate them by comparing different results. For this we use the normalised root-mean square difference (NRMSD) between two responses $R_1$ and $R_2$, given by
%
\begin{equation}
  \mr{NRMSD~(\%)} = 200\,\frac{|R_1 - R_2|}{|R_1| + |R_2|}\ ,
  \label{eq:nrmsd}
\end{equation}
%
to which we refer as simply the \emph{normalised difference} throughout the paper.

The results for the three receiver lines $y=-3;0;+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the normalised differences between the absolute responses of the codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-block}
  \caption{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the normalised differences (\%) between the absolute responses $|E_x|$ of the different codes are shown in the bottom row.}
  \label{fig:results-block}
\end{figure*}
%

A few interesting points stand out:
\begin{enumerate}
  \item The error levels between the different codes are overall comparable; the normalised difference is generally below a few percent, increasing towards larger offsets.
  \item The normalised difference between the FV codes \emg3d and \simpeg is comparably low except at the boundary. The generally low normalised difference is because these codes use the same rectilinear mesh for modelling. The increasing normalised difference towards the boundary is because \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition in this example.
  \item The FE codes \custem and \petgem show similar boundary effects, but are influenced by the anomaly response (higher errors on right side and now unsymmetrical error distribution).
  \item The agreement within the same differential equation method (FE or FV) is better than across methods.
\end{enumerate}

@@

The responses for all six frequencies and all three electric components are shown in Figure~\ref{fig:results-marlim-responses}, both inline and broadside. The published responses are shown in colour with markers for the frequency, and the responses from our codes are shown beneath in grey colours in this overview plot.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim-responses}
  \caption{MR3D comparison between our codes (grey lines) and the published data (coloured lines). The grey lines under the coloured lines are not visible in most areas, meaning that they are very similar. However, there are three notable zones, (1) to (3), which are explained further in the text.}
  \label{fig:results-marlim-responses}
\end{figure*}
%
Two CSEM data-sets were published with MR3D, one containing clean responses, and one responses where realistic noise was added. For the comparison we use the clean data, but we indicate the chosen noise level by the horizontal line at \num{2e-15}\,V/m. It can be seen that the grey lines are not visible for most parts, which means that the data agree well. However, there are a few notable points where this is not the case, and they are marked with numbers in the figure.
\begin{enumerate}
  \item There are some noticeable differences for positive offsets of the broadside $E_y$ and $E_z$ components, which are likely related to the bathymetry. The reason why we cannot reproduce the published results exactly is that the code is not open-source, and we do not know in detail what it does internally. See the discussion around Figure~\ref{fig:results-marlim_2published}.
  \item The three highest frequencies become noisy at large offsets for all of our codes (well below any real-world noise level). The reason why the published responses are not noisy is probably because they were computed in the time domain, and transformed to the frequency domain. This explains their smooth behaviour, without saying anything about their accuracy.
  \item The responses for the $E_y$ component of the inline acquisition line do not agree at all. In the 1D case, the inline $E_y$ component would be zero, and the only response we can measure here comes from 3D effects. These responses are very low, roughly two orders of magnitude lower than the $E_x$ component. Not the published nor our responses are stable, and any of the responses is as bad as the other. Tiniest differences in meshing or different interpolation algorithms will have a huge effect here.
\end{enumerate}

It is important to note that the published results nor any of our results is the \emph{thruth}, but the closer they agree with each other the more certain we can be of the outcomes. Figure~\ref{fig:results-marlim_2published} shows the normalised differences between the published results and our codes for the three strongest components, the inline $E_x$ and the broadside $E_x$ and $E_y$ components, for three frequencies.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2published}
  \caption{Normalised differences for all four codes in comparison with the published data, for the inline $E_x$ field and the broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively. Shown are the three frequencies 0.125\,Hz, 0.5\,Hz, and 1.0\,Hz in the top, middle, and bottom row, respectively.}
  \label{fig:results-marlim_2published}
\end{figure*}
%
A few conclusions can be drawn from this figure: In general the normalised difference is roughly 10\,\% or less. The structured meshes have in some cases a smaller normalised difference, e.g., for negative offsets in the inline and broadside $E_x$ responses, and in other cases the unstructured tetrahedral meshes, e.g., for positive offsets for all broadside $E_y$ responses, and often they have a comparable NRMSD. Some normalised differences have an interesting step-pattern, which is more pronounced for the structured meshes than for the unstructured meshes; this is related to the annotated point (3) in Figure~\ref{fig:model-marlim}. The origin of this zigzag pattern is most likely the bathymetry. We suspect that the code of the published data did probably use the bathymetry information in addition to the resistivity cube. This would explain why the codes using the tetrahedra mesh, which uses the bathymetry as well, does not show this feature.

The importance of the meshing to the result can be seen in Figure~\ref{fig:results-marlim_2ours}, where we compute the normalised differences between our codes.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_2ours}
  \caption{Normalised differences, just as in Figure~\ref{fig:results-marlim_2published}, but comparing some of the four codes with each other. \custem and \petgem produce very similar results, which is due to the fact that they use the same mesh. This also shows that the biggest impact in 3D forward modelling comes from the discretization.}
  \label{fig:results-marlim_2ours}
\end{figure*}
%
The most obvious result in that figure is that \custem and \petgem produce results that are almost identical, their normalised difference is generally below 0.03\,\%. The reason is simple: \custem and \petgem use the exactly same mesh for the computation of this example, the only difference is therefore the solver. This exemplifies that the biggest difference between different codes is the discretization; the solver is important when it comes to runtime and memory, but not for the accuracy of the result. As these two are so similar we compare \emg3d and \simpeg only to \custem in Figure~\ref{fig:results-marlim_2ours}, as the comparisons to \petgem would look the same. But in general the normalised differences between our codes look similar as in the comparison with the published results, and the principal source of the difference must be in the different discretizations.

The actual differences in meshing is shown in the next two figures. Figure~\ref{fig:results-marlim_big} shows on the top row the rectilinear mesh of \emg3d, on the middle row the octree mesh of \simpeg, and on the bottom row the tetrahedral mesh of \custem and \petgem. The actual mesh and the broadside $|E_x|$ and $|E_y|$ fields are shown in the left, middle, and right columns, respectively. The extent of the plot includes the entire computational domain for the tetrahedral mesh (in x and z direction); the rectilinear and octree mesh extents are larger than what is shown (rectilinear: $x=327$ to 454\,km, $z=-57$ to 66\,km; octree: $x=288$ to 492\,km, $z=-123$ to 41\,km).
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_big}
  \caption{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively. The shown dimensions shows the entire computational domain for the tetrahedra mesh, whereas the entire rectilinear and octree meshes are bigger.}
  \label{fig:results-marlim_big}
\end{figure*}
%
This figure is interesting for various points, as the differences between the three meshes are clearly visible. Particularly the advantage of octree and tetrahedra to create large cells outside of the region of interest to push away the boundary can be seen, whereas a rectilinear grid, even though stretched, has to compute many cells far away of the zone of interest. The different boundary condition also play a part, where the top row used PEC, and the other two rows PMC.

Figure~\ref{fig:results-marlim_survey} shows a zoom-in to the survey domain, the domain of interest, which shows how the different codes discretized the zones of interest.
%
\begin{figure*}
  \centering
  \includegraphics[width=.9\linewidth]{figures/results-marlim_survey}
  \caption{Meshes and broadside $E_x$ and $E_y$ fields in the left, middle, and right column, respectively, for the rectilinear, octree, and tetrahedral meshes in the top, middle, and bottom row, respectively (3$\times$ vertical exaggeration).}
  \label{fig:results-marlim_survey}
\end{figure*}
%
The most striking point of these mesh comparison plots is that, while the overall fields are similar and the responses at receiver locations agree well, the fields can vary quite a lot in different parts of the model. Two kilometres below the seafloor, as an example, there are already significant changes visible, purely due to meshing. This is an important insight that has to be taken into account when interpreting modelling results and the feasibility for inversions.

@@

%% 1. Results: good agreement %%
Overall, we observed an excellent match between the solutions. The amplitudes of the layered model have mostly a relative error of less than 1--2\,\%. Looking individually at the real or imaginary part results in much higher error values, but those are related to the zero crossings of the responses. The cross-validation of the results for the block model yields a similar picture, with normalised differences within a few percents. For the large, complex MR3D model it is a bit different. First it can be noted that the misfits are higher for higher frequencies. For the lowest frequency, 0.125\,Hz, the normalised difference is a few percent for inline and broadside $E_x$ components, and up to 10\,\% for the broadside $E_y$ component. For higher frequencies these numbers rise to the order of $\pm10\,$\%. The reason for the higher misfit for MR3D than for the layered or block model lies mostly in the discretization, boundary effects, and interpolation strategy. While one might argue that 5--10\,\% NRMSD is a lot, we argue that this is probably as good as it gets, given that the code with which the responses we compare to were computed is not open-source. This has been shown within the process of this study. Our first attempts were based on the originally published, fine MR3D model. This model is very detailed, but its extent is too small in x- and y-directions for CSEM computations. In order to represent the fine-scale model on a mesh suitable for computation, each of our codes did its own upscaling and extrapolation. However, we were unsuccessful in our attempt to reproduce the published results.  The reason is that there are many ways how you can extrapolate a model, and without knowing the details of how MR3D represents the model on the computational mesh, we could not re-produce this step. We reached out to the authors of MR3D, and they kindly agreed to also published the upscaled and extended computational model. It is only with this model that we were able to obtain comparable results.

%% 2. Discretization %%
One of the key insights we gained from this study is that creating appropriate meshes is a difficult and time-consuming task. While there now exist several open-source 3D CSEM codes the capability for automatic meshing lacks behind and currently remains a largely manual task requiring experience in the field. It is important to state that the chosen models favour structured meshes. The first two examples can easily and accurately be represented by regular meshes. Using, for instance, a dipping layer instead of three resistive cubes would turn it around and make the model more favourable for unstructured meshes. MR3D would be an ideal scenario for unstructured meshes, as it is based on also published irregular lithological horizons. Since the corresponding resistivity model, however, was defined on a regular grid, the FE codes were forced to approximate the comparatively fine regular discretization by an unstructured one and interpolate the resistivity data for being able to apply the FE codes to this problem. We are completely aware about the ineptness of this re-approximation procedure, unless it served for the cross-validation purposes. Both cases show that automatic and adaptive meshing capabilities are needed, which will increase the flexibility of 3D CSEM modelling. Such meshing codes should also take into account the physics of CSEM computation with its diffusive behaviour. It is also important to state that the unstructured mesh used for the MR3D model incorporates the entire domain and could be used for all source and receiver positions, whereas the rectilinear and octree meshes were designed for the shown receiver and corresponding source lines.

@@

The resulting CSEM responses of the four codes for the three models, as well as all files to rerun the different models with the four codes and reproduce the shown results are available at Zenodo (DOI) \emph{with the final submission}. For the review process we refer to the github page \href{https://github.com/swung-research/3d-csem-open-source-landscape}{github.com/swung-research/3d-csem-open-source-landscape}.
